Measuring causal influence with\\ back-to-back regression: the linear case

### 2019-05-23 paper/meg_exps.tex
back to 8 pages

\subsubsection{MEG preprocessing}

The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies.

The raw data was then segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz. Figure 1 in the supplementary material
shows the average response of the magnetometers evoked by the words presented
to the first subject.

For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 2700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

\subsubsection{Word features} We annotated each word with 38 features: 26
correspond to the count of each letter in the word (from 'a' to 'z',
independently of their case), the total number of letters ('word length'), the
frequency of the word in natural language (using the Zipf logarithmic scale
of \citep{van2014subtlex}), and the part-of-speech tags (i.e. NOUN, PROPN (proper
noun), VERB, PRON (pronoun), CONJ (conjunction), DET (determinant), ADJ
(adjective), ADV (adverb), and NUM (numerical)) using Spacy \citep{spacy2} and
WordFreq \citep{speerwordfreq}. For clarity purposes, we refer to these
features using two broad categories: visual (i.e. each letter and the total
number of letters) and lexical (part-of-speech and word frequency). This
procedure yields an $X \in \mathbb{R}^{n \times d_x}$ matrix of $n\approx$ 2700 words by
$d_x=38$ features for each subject. Each of the columns of $X$ is normalized to
have a mean and a standard deviation of 0 and 1 respectively.

\subsubsection{Methods}

@@

Forward regression was performed with an $\ell_2$-regularized least-squares regression $H \in \mathbb{R}^{d_x \times d_y}$ from $X$ to $Y$, and estimates the causal influence $E_{i,i}$ as $\| H_{i, :} \|^2$.
%
Backward regression was performed with an $\ell_2$-regularized least-squares regression $G \in \mathbb{R}^{d_y \times d_x}$ from $Y$ to $X$, and estimates the causal influence $E_{i, i}$ as the Pearson correlation between the $i$-th column of $X$ and the $i$-th column of $YG$, using a validation sample.
%
Finally, we used the back-to-back regression was performed using two $\ell_2$-regularized
least squares regressions $G$ and $H$ as described in section \ref{sec:algorithm}, using a bagging factor of 10, and alpha cross-validated between $10^{-5}$ and $10^{5}$.

\subsubsection{Second-level statistics}

>>>>

\subsubsection{MEG preprocessing and feature generation}

The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies. The raw data was then segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz. Figure 1 in the supplementary material
shows the average response of the magnetometers evoked by the words presented
to the first subject. For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 2700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

We annotated each word with 38 features: 26
correspond to the count of each letter in the word (from 'a' to 'z',
independently of their case), the total number of letters ('word length'), the
frequency of the word in natural language (using the Zipf logarithmic scale
of \citep{van2014subtlex}), and the part-of-speech tags (i.e. NOUN, PROPN (proper
noun), VERB, PRON (pronoun), CONJ (conjunction), DET (determinant), ADJ
(adjective), ADV (adverb), and NUM (numerical)) using Spacy \citep{spacy2} and
WordFreq \citep{speerwordfreq}. For clarity purposes, we refer to these
features using two broad categories: visual (i.e. each letter and the total
number of letters) and lexical (part-of-speech and word frequency). This
procedure yields an $X \in \mathbb{R}^{n \times d_x}$ matrix of $n\approx$ 2700 words by
$d_x=38$ features for each subject. Each of the columns of $X$ is normalized to
have a mean and a standard deviation of 0 and 1 respectively.

\subsubsection{Models and statistics}

@@

Forward regression was performed with an $\ell_2$-regularized least-squares regression $H \in \mathbb{R}^{d_x \times d_y}$ from $X$ to $Y$, and estimates the causal influence $E_{i,i}$ as $\| H_{i, :} \|^2$.
Backward regression was performed with an $\ell_2$-regularized least-squares regression $G \in \mathbb{R}^{d_y \times d_x}$ from $Y$ to $X$, and estimates the causal influence $E_{i, i}$ as the Pearson correlation between the $i$-th column of $X$ and the $i$-th column of $YG$, using a validation sample. Finally, we used the back-to-back regression was performed using two $\ell_2$-regularized
least squares regressions $G$ and $H$ as described in section \ref{sec:algorithm}, using a bagging factor of 10, and alpha cross-validated between $10^{-5}$ and $10^{5}$.
